---
title: "QAQC - SITES"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
output:
  word_document:
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: 3
mainfont: Cambria
geometry: margin=.5in
classoption: portrait
---
```{r include=FALSE}
library(knitr)
library(tidyverse)
library(scales)

QAQCSitesOut <- readRDS("../Temp_out/QAQCSitesOut.RDS")

FuncPlotPSA <- function(PSA_df, xlabel) {
  # Function to create PSA plot
  # 
  # Args:
  #   PSA_df:  Dataframe output from FuncPSA
  #   xlabel: Text for x-axis label (category)
  #
  # Returns:
  #   PSA point plot with bootstrapped 90% CI's, from within function
  ggplot(PSA_df, aes(x=Grp,  y=PSA)) +
    geom_point(na.rm = TRUE) +
    geom_linerange(aes(ymin=`5%`, ymax=`95%`), na.rm = TRUE) +
    ylim(0, 100) +
    scale_x_discrete(labels=paste0(PSA_df$Grp,"\n(",PSA_df$Denom,")")) +
    labs(y="Proportion of specific agreement (%)", x=xlabel, subtitle="(90% CI included only if 5 or more positive detections)") +
    theme_bw(base_size = 10)
}

FuncBarHist <- function (dat, plot_vals, alt_xlab = NA) {
  # Function to generate a histogram of count/continuous data differences between QAQC and non-QAQC surveys 
  # 
  # Args:
  #   dat:  Dataframe of data
  #   plot_vals:  Vector of plot values
  #   alt_xlab:  Alternative text for the x-axis label
  #
  # Returns:
  #   Generates a bar plot or histogram from within function
  #
  p <- ggplot(data = dat, aes(x = Metric)) +
  scale_x_continuous(limits = c(-1,1)*plot_vals$range_max, breaks = seq(-1*plot_vals$range_max, plot_vals$range_max, by = plot_vals$range_int)) +
  theme_bw(base_size = 12)
  
  if (plot_vals$type == "discrete") { # if count data, return a bar plot
    p + geom_bar(fill = "white", color = "black", na.rm = TRUE) +
      geom_vline(xintercept = 0, col = "red") +
      labs(x = ifelse(is.na(alt_xlab), "Difference between surveys, q - n", alt_xlab), y = "Count")
  } else {
    p + geom_histogram(fill = "white", color = "black", binwidth = plot_vals$range_int, na.rm = TRUE) +
      geom_vline(xintercept = 0, col = "red") +
      labs(x = ifelse(is.na(alt_xlab), "% difference between surveys, (q - n)/mean", alt_xlab), y = "Count")
  }
}

FuncBAPlot <- function(dat, plot_vals, alt_ylab = NA) {
  # Function to generate a Bland-Altman plot 
  # 
  # Args:
  #   dat:  Dataframe of data
  #   plot_vals:  Vector of plot values
  #   alt_ylab:  Alternative text for the y-axis label
  #
  # Returns:
  #   Generates a Bland-Altman plot from within function
  #
  p <- ggplot(data=dat, aes(x=Mean, y=Metric)) +
  geom_point(alpha = 0.20, na.rm = TRUE) +
  stat_smooth(method="loess") +
  geom_hline(yintercept=quantile(dat$Metric, probs=c(0.025, 0.975)), col="blue", linetype="dashed") +
  geom_hline(yintercept = 0, col="red") +
  scale_y_continuous(breaks=pretty_breaks()) +
  theme_bw(base_size = 12)
  
  if (plot_vals$type == "discrete") { # if count data
    p + labs(x="Survey mean", y = ifelse(is.na(alt_ylab), "Difference between surveys, q - n", alt_ylab), subtitle="(gray shading is 95% CI for loess curve; \ndashed blue lines span 95% of the difference values)")
    } else { # if continuous
      p + labs(x="Survey mean", y = ifelse(is.na(alt_ylab), "Difference between surveys, q - n, as % of mean", alt_ylab), subtitle="(gray shading is 95% CI for loess curve; \ndashed blue lines span 95% of the difference values)")
    }
}

```

\newpage

### COMMUNITY CODE & COMMUNITY GROUP ASSIGNMENT

Data Notes:

* For each classification, there is also a 'fit', where 1=confident and 2,3=not confident about the classification

* This is a required field

* NA/NA is a MISMATCH

&nbsp;

**Community Code summary**
*In the row headings, 1 means fit=1; + means fit=2 or 3. In the column headings, FALSE means Community Code (CC) did not match between the two surveys and TRUE means CC did match. The numbers in cells are numberes of QAQC pairs.*
```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
CG_df <- QAQCSitesOut[["CG_list"]][["GroupSummary"]]
kable(table(CG_df$FitMatch, CG_df$CCMatch))
```

**Community Group summary**
*Same as above, but for Community Group instead of Community Code.*
```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
kable(table(CG_df$FitMatch, CG_df$CGMatch))
```

**Community Group PSA**
*Proportion of specific agreement. For example, for Community Group G031... there were 3 QAQC pairs in which the site was classified as G031 by BOTH the QAQC surveyor and the non-QAQC surveyor (numerator=2 X number of matches for that CG). There were 7 instances total in which a site was classified as G031 by either the QAQC surveyor or the non-QAQC surveyor.*
```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
kable(QAQCSitesOut[["CG_list"]][["CG_PSA_df"]])
```

**Community Group PSA plot**
```{r fig.width=7.5, fig.height=3.5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
FuncPlotPSA(PSA_df = QAQCSitesOut[["CG_list"]][["CG_PSA_df"]], xlabel = "Community Group")
```

\newpage

### STAND HEIGHT

Data Notes:

* This metric was added to the protocol late

* NA/NA is OMITTED

* Summary excluded plot-surveys in which only one survey recorded a stand height

&nbsp;

**Cumulative distribution table for absolute % differences in stand height, (q - n)/mean**
*Table should be read as, e.g., X% of stands (second column) have an absolute % difference in recorded stand height of Y% or less (first column).*
```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
kable(QAQCSitesOut[["StandHeight"]][["Diff_tab"]], col.names=c("Max. absolute % difference", "Cumulative % of stands"))
```

**Histogram of % differences in stand height, (q - n)/mean**
```{r fig.width=6.5, fig.height=3.5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
FuncBarHist(dat = QAQCSitesOut[["StandHeight"]][["FinalDat_df"]], plot_vals = QAQCSitesOut[["StandHeight"]][["plot_vals"]])
```

\newpage

**Stand height absolute % differences, empirical cumulative distribution function**

*Vertical red lines at 5% & 10%*
```{r fig.width=4.5, fig.height=4.5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
plot.ecdf(abs(QAQCSitesOut[["StandHeight"]][["FinalDat_df"]][["Metric"]]), col.01line="white", main = "CDF for % difference in stand height", ylab="Proportion of stands", xlab="% Difference in stand height", cex.lab=1, cex.main=1)
abline(v=5, col="red", lty="dashed")
abline(v=10, col="red", lty="dashed")
```

**Bland-Altman plot for stand height**

*Darker points represent more data values*
```{r fig.width=5.5, fig.height=4.5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
FuncBAPlot(dat = QAQCSitesOut[["StandHeight"]][["FinalDat_df"]], plot_vals = QAQCSitesOut[["StandHeight"]][["plot_vals"]])
```

### SITE DISTURBANCE

Data Notes: 

* The datasheet had a checkbox for each disturbance, so this is essentially a required field (unchecked means FALSE; checked means TRUE for each disturbance type)

* NA/NA is a MISMATCH

&nbsp;

**Site disturbance summary & PSA**
```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
kable(QAQCSitesOut[["SiteDisturb_list"]][["SiteDisturbPSASummary_df"]])
```

**Site disturbance PSA plot**
```{r fig.width=7.5, fig.height=3.5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
QAQCSitesOut[["SiteDisturb_list"]][["SiteDisturbPSA_plot"]]
```

### COARSE WOODY DEBRIS COUNT BY PLOT

Data Notes:

* If transect information (x or y) is missing, data are used for plot-level counts, but not included for transect-level counts because no information on which transect to assign counts to

**Frequency table for absolute differences in number of CWD by plot**
```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
kable(QAQCSitesOut[["CWDPlotCounts"]][["Diff_tab"]], col.names=c("Difference Category", "Freq"))
```

**Histogram of differences in CWD counts by plot**
```{r fig.width=6.5, fig.height=3.5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
FuncBarHist(dat = QAQCSitesOut[["CWDPlotCounts"]][["FinalDat_df"]], plot_vals = QAQCSitesOut[["CWDPlotCounts"]][["plot_vals"]])
```

\newpage

**Absolute differences in CWD counts by plot, empirical cumulative distribution function**

*Vertical red lines at 0 & 1*
```{r fig.width=4.5, fig.height=4.5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
plot.ecdf(abs(QAQCSitesOut[["CWDPlotCounts"]][["FinalDat_df"]][["Metric"]]), col.01line="white", main = "CDF for difference in CWD count", ylab="Proportion of plots", xlab="Absolute difference in CWD count", cex.lab=1, cex.main=1)
abline(v=0, col="red", lty="dashed")
abline(v=1, col="red", lty="dashed")
```

**Bland-Altman plot for CWD counts by plot**

*Darker points represent more data values*
```{r fig.width=5.5, fig.height=4.5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
FuncBAPlot(dat = QAQCSitesOut[["CWDPlotCounts"]][["FinalDat_df"]], plot_vals = QAQCSitesOut[["CWDPlotCounts"]][["plot_vals"]])
```

### COARSE WOODY DEBRIS COUNT BY TRANSECT

**Frequency table for absolute differences in number of CWD by transect**
```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
kable(QAQCSitesOut[["CWDTransectCounts"]][["Diff_tab"]], col.names=c("Difference Category", "Freq"))
```

**Histogram of differences in CWD counts by transect**
```{r fig.width=6.5, fig.height=3.5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
FuncBarHist(dat = QAQCSitesOut[["CWDTransectCounts"]][["FinalDat_df"]], plot_vals = QAQCSitesOut[["CWDTransectCounts"]][["plot_vals"]])
```

\newpage

**Absolute differences in CWD counts by transect, empirical cumulative distribution function**

*Vertical red lines at 0 & 1*
```{r fig.width=4.5, fig.height=4.5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
plot.ecdf(abs(QAQCSitesOut[["CWDTransectCounts"]][["FinalDat_df"]][["Metric"]]), col.01line="white", main = "CDF for difference in CWD count", ylab="Proportion of transects", xlab="Absolute difference in CWD count", cex.lab=1, cex.main=1)
abline(v=0, col="red", lty="dashed")
abline(v=1, col="red", lty="dashed")
```

**Bland-Altman plot for CWD counts by transect**

*Darker points represent more data values*
```{r fig.width=5.5, fig.height=4.5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
FuncBAPlot(dat = QAQCSitesOut[["CWDTransectCounts"]][["FinalDat_df"]], plot_vals = QAQCSitesOut[["CWDTransectCounts"]][["plot_vals"]])
```

**Distribution of decay class**

```{r fig.width=4, fig.height=4, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
CWD <- QAQCSitesOut[["CWDPlotCounts"]][["CWD_orig_df"]]
CWD$QAQC <- factor(CWD$QAQC, labels=c("Nsurv", "Qsurv"))
ggplot(CWD) +
  geom_bar(aes(x = Decay)) +
  theme_bw(base_size = 10) +
  labs(x="Decay class", y="Number of CWD") +
  facet_grid(QAQC ~.)
```

**Distribution of diameter**

```{r fig.width=4, fig.height=4, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
ggplot(CWD) +
  geom_histogram(aes(x = Diameter), binwidth = 3) +
  theme_bw(base_size = 10) +
  labs(x="Diameter", y ="Number of CWD") +
  facet_grid(QAQC ~.)
```