---
title: "QAQC - HERBS - COVER CLASS & SPECIES RICHNESS (`r params$type`)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
mainfont: Cambria
geometry: margin=.5in
params:
  type: HerbsBase
classoption: landscape
---
```{r include=FALSE}
library(knitr)
library(ggplot2)
library(cowplot)

QAQCHerbsOut <- readRDS("../Temp_out/QAQCHerbsOut.RDS")
QAQCHerbsDat <- readRDS("../Temp_out/QAQCHerbsDat.RDS")

FuncHerbsCover <- function (Dat_df) {
  # Function to summarize herb match types. Use only for 100sqm-level analysis, and all species or hicov (not rolled to genus) 
  # 
  # Args:
  #   Dat_df:  A data frame of the cleaned herb data
  # Returns:
  #   A list object that can be printed via an .RMD file
  HerbsCC_df <- Dat_df
  HerbsCC_df$CC[is.na(HerbsCC_df$CC)] <- 0 # Zero for now
  HerbsCC_df <- HerbsCC_df %>%
    select(PlotID, QAQC, PlantName, CC) %>%
    spread(key=QAQC, value=CC)
  HerbsCC_df <- HerbsCC_df[complete.cases(HerbsCC_df),]
  colnames(HerbsCC_df)[3:4] <- c("Nsurv", "Qsurv")
  HerbsCC_df$Nsurv[HerbsCC_df$Nsurv==0] <- NA # convert back to NA
  HerbsCC_df$Qsurv[HerbsCC_df$Qsurv==0] <- NA
  HerbsCC_df$AbsDiff <- abs(HerbsCC_df$Qsurv - HerbsCC_df$Nsurv)
  HerbsCC_df <- as.data.frame(HerbsCC_df)
  return(HerbsCC_df)
}

FuncPlotHerbsRichUnmatched <- function (HM_df, MinDetect, ScaleLab) {
  # Function to generate a bar plot of proportion of species detections missed in QAQC surveys.
  # 
  # Args:
  #   HM_df:  Herb match data frame.
  #   MinDetect:  Minimum number of detections for a species to be included
  #   ScaleLab:  Character string indicating the scale of evaluation (options are "Plot", "Module", "Eighth")
  # Returns:
  #   Bar plot
  #
  HFspecies <- which(table(HM_df$PlantName) >= MinDetect) # species detected on at least X plot/mod/eighth-surveys
  HFmissed <- HM_df %>%
    filter(PlantName %in% names(HFspecies)) %>%
    group_by(PlantName) %>%
    dplyr::summarize(Count = n(),
              AbsMissed = Count - sum(MatchTT),
              PropMissed = as.integer(round((sum(!MatchTT)/Count) * 100))) %>%
    mutate(PlantName = paste0(PlantName, " (", Count, ")"))
  
  BarMissed_plot <- ggplot() +
    geom_bar(data = subset(HFmissed, PropMissed > 50), aes(x=reorder(PlantName, -PropMissed), y = PropMissed), width=.5, stat="identity") +
    theme_bw(base_size=10) +
    ylim(0, 100) +
    geom_hline(yintercept=50, color="red") +
    labs(y="Proportion of Detections Unmatched (%)", x="Species", title= paste0("Proportion of ", ScaleLab, "Survey Detections Unmatched"), subtitle=paste0("(Species with at least ", MinDetect, " detections and >50% unmatched)")) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5)) +
    theme(plot.title = element_text(size=10), legend.title = element_blank())
  
  return(BarMissed_plot)
}

FuncPlotHerbsRichDiff <- function (HS_df, ScaleLab) {
  # Function to generate a bar plot of proportion of species detections missed in QAQC surveys.
  # 
  # Args:
  #   HM_df:  Herb match data frame.
  #   MinDetect:  Minimum number of detections for a species to be included
  #   ScaleLab:  Character string indicating the scale of evaluation (options are "Plot", "Module", "Eighth")
  # Returns:
  #   Histogram and scatterplot
  #
  HistRichDiff_plot <- 
    ggplot(HS_df, aes(x=RichDiff)) +
    geom_bar() +
    geom_vline(xintercept=0, col="red") +
    labs(subtitle="(Positive difference means QAQC higher)", x=paste0("Difference in Herb ", ScaleLab, " Richness"),  y=paste0("Number of ", ScaleLab, "-Surveys")) +
    theme_bw(base_size = 12)
  
  PointsRichDiff_plot <-
    ggplot(HS_df, aes(x=NsurvSpp, y=QsurvSpp)) +
    geom_point() +
    geom_abline(slope=1, col="red") +
    labs(x=paste0("Herb ", ScaleLab, " Richness in Normal Survey"), y=paste0("Herb ", ScaleLab, " Richness in QAQC Survey")) +
    theme_bw(base_size = 12)
  
  return_list <- list(HistRichDiff_plot, PointsRichDiff_plot)
  return(return_list)
}
```

\newpage

### HERB COVER CLASS

Data Notes:

* Herb cover class is evaluated only at the 100sq.m. level.

* Only use species-100sq.m.-surveys with detection in both the QAQC and non-QAQC survey.

* 9 cover classes of different spans.

&nbsp;

**Herbs Cover Class Empirical Cumulative Distribution Function**

&nbsp;

```{r echo=FALSE, results='as is'}
if(params$type == "HerbsRolled") {
  cat("Cover class summaries not applicable for data with some species rolled to genus level")
}
```

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
if(params$type %in% c("HerbsBase", "HerbsHiCov")) {
  HerbsDat <- QAQCHerbsDat[[params$type]]
  HerbsCC_df <- FuncHerbsCover(HerbsDat)
  
  HS_tab <- data.frame(table(factor(HerbsCC_df$AbsDiff, levels = min(HerbsCC_df$AbsDiff, na.rm = TRUE):max(HerbsCC_df$AbsDiff, na.rm = TRUE))))
  HS_tab$CDF <- round(cumsum(HS_tab$Freq)/sum(HS_tab$Freq)*100, 1)
  names(HS_tab) <- c("Difference in Counts", "Frequency", "CDF (%)")
  kable(HS_tab)
}
```

\newpage

```{r fig.width=7, fig.height=7, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
if(params$type %in% c("HerbsBase", "HerbsHiCov")) {
  QAQC_df <- HerbsCC_df[, c("Nsurv", "Qsurv")]
  QAQC_df[is.na(QAQC_df)] <- "XXX"
  
  categ_vec <- c(1:9, "XXX")
  QAQC_df[, 1] <- as.integer(factor(QAQC_df[, 1], levels=categ_vec)) # convert categories to numbers
  QAQC_df[, 2] <- 
    as.numeric(factor(QAQC_df[, 2], levels=categ_vec))
  
  QAQC_mat <- as.matrix(QAQC_df)
  QAQC_mat <- t(apply(QAQC_mat,1,sort)) # For upper triangular, within each row, the smaller number is first
  QAQC_squaremat <- 
    table(factor(QAQC_mat[,1], levels=1:length(categ_vec)),
          factor(QAQC_mat[,2], levels=1:length(categ_vec)))
  
  oldpar <- par(no.readonly = TRUE)
  par(oma=c(0, 0, 0, 0))
  par(cex=.8)
  par(mar=c(3.1, 4.1, 7.1, 3.1))
  
  cellcol <-
    color.scale(cbind(QAQC_squaremat, seq(min(QAQC_squaremat, na.rm=TRUE), max(QAQC_squaremat, na.rm=TRUE), length.out=nrow(QAQC_squaremat))), c(1,0,0), c(1,1,0), c(0,1,1))[, 1:dim(QAQC_squaremat)[2]]
cellcol[QAQC_squaremat==0] <- "white"
cellcol[lower.tri(cellcol)] <- "black"
QAQC_squaremat[lower.tri(QAQC_squaremat)] <- NA
(color2D.matplot(QAQC_squaremat, cellcolors=cellcol, show.legend=FALSE, show.values=TRUE, axes=FALSE, xlab="", ylab=""))
(axis(3, at=0.75:dim(QAQC_squaremat)[2]-.25, tick=FALSE, labels=categ_vec, las=2))
(axis(2, at=.75:dim(QAQC_squaremat)[1]-.25, tick=FALSE, las=2, labels=rev(categ_vec)))
title(main=paste0("QAQC Evaluation - Herbs CC (N = ", sum(QAQC_squaremat, na.rm=TRUE), ")"), line=4.4)
par(oldpar)  # set margins back to default
}
```

\newpage

### HERB RICHNESS & SPECIES - 100SQ.M. LEVEL

**Herbs 100sq.m. richness summary**

&nbsp;

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
HerbsPlotSummary <- QAQCHerbsOut$HS100SqM_list[[params$type]]
kable(summary(HerbsPlotSummary[c("NsurvSpp", "QsurvSpp", "PSturn", "PercRichDiff", "UniqueSpp", "MatchTrue", "MatchByMean", "MatchByUnique")]))
```

\newpage

**Differences in herb richness at 100sq.m. level**

&nbsp;

```{r fig.width=9, fig.height=5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
HerbsDiffPlots_list <- FuncPlotHerbsRichDiff(HS_df = HerbsPlotSummary, ScaleLab = "Plot")
plot_grid(HerbsDiffPlots_list[[1]], HerbsDiffPlots_list[[2]], ncol=2)
```

\newpage

**Herb Richness Empirical Cumulative Distribution Function (100sq.m.)**

&nbsp;

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
kable(QAQCHerbsOut$HS100SqMTable_list[[params$type]], align = "c")
```

\newpage

**Herb species most missed at 100sq.m. level**
```{r fig.width=9, fig.height=6, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
(FuncPlotHerbsRichUnmatched(HM_df = QAQCHerbsOut$HM100SqM_list[[params$type]], MinDetect = 5, ScaleLab = "Plot"))
```  

\newpage

### HERB RICHNESS & SPECIES - 10SQ.M. LEVEL

**Herbs 10sq.m. richness summary**

&nbsp;

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
HerbsModSummary <- QAQCHerbsOut$HS10SqM_list[[params$type]]
kable(summary(HerbsModSummary[c("NsurvSpp", "QsurvSpp", "PSturn", "PercRichDiff", "UniqueSpp", "MatchTrue", "MatchByMean", "MatchByUnique")]))
```

\newpage

**Differences in herb richness at 10sq.m. level**

&nbsp;

```{r fig.width=9, fig.height=5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
HerbsDiffMod_list <- FuncPlotHerbsRichDiff(HS_df = HerbsModSummary, ScaleLab = "Mod")
plot_grid(HerbsDiffMod_list[[1]], HerbsDiffMod_list[[2]], ncol=2)
```

\newpage

**Herb Richness Empirical Cumulative Distribution Function (10sq.m.)**

&nbsp;

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
kable(QAQCHerbsOut$HS10SqMTable_list[[params$type]], align = "c")
```

\newpage

**Herb species most missed at 10sq.m. level**
```{r fig.width=9, fig.height=6, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
(FuncPlotHerbsRichUnmatched(HM_df = QAQCHerbsOut$HM10SqM_list[[params$type]], MinDetect = 30, ScaleLab = "Mod"))
```  

\newpage

### HERB RICHNESS & SPECIES - 1SQ.M. LEVEL

**Herbs 1sq.m. richness summary**

&nbsp;

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
HerbsEighthSummary <- QAQCHerbsOut$HS1SqM_list[[params$type]]
kable(summary(HerbsEighthSummary[c("NsurvSpp", "QsurvSpp", "PSturn", "PercRichDiff", "UniqueSpp", "MatchTrue", "MatchByMean", "MatchByUnique")]))
```

\newpage

**Differences in herb richness at 1sq.m. level**

&nbsp;

```{r fig.width=9, fig.height=5, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
HerbsDiffEighth_list <- FuncPlotHerbsRichDiff(HS_df = HerbsEighthSummary, ScaleLab = "Eighth")
plot_grid(HerbsDiffEighth_list[[1]], HerbsDiffEighth_list[[2]], ncol=2)
```

\newpage

**Herb Richness Empirical Cumulative Distribution Function (1sq.m.)**

&nbsp;

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
kable(QAQCHerbsOut$HS1SqMTable_list[[params$type]], align = "c")
```

\newpage

**Herb species most missed at 1sq.m. level**
```{r fig.width=9, fig.height=6, fig.align='center', results='hide', echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
(FuncPlotHerbsRichUnmatched(HM_df = QAQCHerbsOut$HM1SqM_list[[params$type]], MinDetect = 60, ScaleLab = "Eighth"))
```  
